---
title: "School Culture"
output:
  flexdashboard::flex_dashboard:
    orientation: columns
    vertical_layout: fill
    theme: lumen

runtime: shiny
---

```{r setup, include=FALSE}
library(flexdashboard)
library(Cairo)
library(lubridate)
library(ggplot2)
library(dplyr)
library(tidyr)
library(formattable)
library(forcats)

options(shiny.usecairo = TRUE)

load("/data/attendance.Rda")

load("/data/transfers.Rda")

load("/data/suspensions.Rdata")

load("/data/dl_suspensions.Rda")

load("/data/tb_observations.Rda")

config <- as.data.frame(read.dcf("/config/config.dcf"),
                        stringsAsFactors = FALSE)
#Load observation data
# obsData <- reactiveFileReader(8.64e+7,
#   session = NULL,
#   filePath = "/data/tb_observations.Rda",
#   load,
#   envir = .GlobalEnv)

schools <- unique(ada_weekly_school$schoolabbreviation)

grades <- list(KAP = 0:5,
               KAMS = 6:8,
               KCCP = 5:8,
               KBCP = 5:8,
               KOP = 0:1,
               KOA = 5:6)

month_order <- c("Oct", "Nov", "Dec", "Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sep")               

source('lib/attendance_helpers.R')

# ADA goals for each school
goals <- data_frame(schoolabbreviation =  c("KAP", "KAMS", "KAC/KCCP", "KBCP", "KOP", "KOA"),
                    goal = c(.98))

firstday <- ymd(config$FIRST_DAY)
lastXweeks <- ymd(as.character(floor_date(today() - weeks(6), unit="week")+1))
last6weeks <- as.character(max(lastXweeks, firstday))
thisweek <- as.character(today())
firstweek <- as.character(floor_date(min(attend_date_school_grade$week_of_date)))


#firstweek <- as.character(max(firstweek, firstday))



restart_time<-file.info('restart.txt')$mtime

update_time_stamp <- lubridate::stamp("Attendance data last updated on Tuesday, September 14, 2001 at 4:41 pm")(restart_time)

```


```{r server_attendance, include = FALSE}
att_date_range <- reactive(input$att_dates[[1]] %--% input$att_dates[[2]])



  # create dynamic panel
  output$school_grades_ui <- renderUI({

    school_grades <- ada_weekly_grade_hr %>%
      filter(schoolabbreviation == input$att_school) %>%
      select(grade_level)

    grades <- school_grades$grade_level %>% unique()

    selectInput("att_grades",
                 "Choose grade",
                 choices = as.character(grades),
                 selected = as.character(grades[1]),
                 multiple = FALSE)
    })

  # create daily data!
  ada_daily <- reactive({

    # Get data at school level
   if(input$att_level == "school") {


    daily_out <- attend_date_school %>%
      select(name = schoolabbreviation,
             date,
             week_in_sy,
             week_of_date,
             week_of_date_short_label,
             enrolled,
             present) %>%
      mutate(goal = round(.97*enrolled),
             day = wday(date)) %>%
      gather(key = "variable",
             value = "value", enrolled:goal)

   }

    # Munge data for Grade level
    # To do: create menu item for school
    if(input$att_level == "grade") {

      daily_out <- attend_date_grade %>%
        filter(schoolabbreviation == input$att_school) %>%
        select(name = grade_level,
               schoolabbreviation,
               date,
               week_in_sy,
               week_of_date,
               week_of_date_short_label,
               enrolled,
               present) %>%
        mutate(goal = round(.97*enrolled),
               day = wday(date)) %>%
        gather(key = "variable",
               value = "value", enrolled:goal)

    }

    if(input$att_level == "hr") {

      daily_out <- attend_date_grade_hr %>%
        filter(grade_level == as.numeric(input$att_grades),
               schoolabbreviation == input$att_school
               ) %>%
        select(name = home_room,
               grade_level,
               schoolabbreviation,
               date,
               week_in_sy,
               week_of_date,
               week_of_date_short_label,
               enrolled,
               present) %>%
        mutate(goal = round(.97*enrolled),
               day = wday(date)) %>%
        gather(key = "variable",
               value = "value", enrolled:goal)

    }

  # return
  daily_out

  })







  # ada_weekly reactive
  ada_weekly <- reactive({

    # if school
    if(input$att_level == "school") {
      weekly_out <- ada_weekly_school %>%
        mutate(name = schoolabbreviation,
               threshold = ifelse(weekly_ada>=.97*100,
                                  sprintf("Weekly ≥ %s%%", round(.97*100)),
                                  sprintf("Weekly < %s%%",  round(.97*100))
                                  )
              )
     }



    # if grade
    if(input$att_level == "grade") {

      weekly_out <- ada_weekly_school_grade %>%
      filter(schoolabbreviation == input$att_school) %>%
        mutate(name = grade_level,
               threshold = ifelse(weekly_ada>=.97*100,
                                  sprintf("Weekly ≥ %s%%", round(.97*100)),
                                  sprintf("Weekly < %s%%",  round(.97*100))
                                  )
              )
     }

    # if home room
    if(input$att_level == "hr") {


      weekly_out <- ada_weekly_grade_hr %>%
        filter(grade_level == as.numeric(input$att_grades),
               schoolabbreviation == input$att_school
               ) %>%
        mutate(name = home_room,
               threshold = ifelse(weekly_ada>=.97*100,
                                  sprintf("Weekly ≥ %s%%", round(.97*100)),
                                  sprintf("Weekly < %s%%",  round(.97*100))
                                  )
              )
   }

  #return
  weekly_out

  })






  # Step plot of enrollment and attendance
  output$step_plot <- renderPlot({

    # verify that initial grade level exists for school if att_level == 'hr'
    # you need to do this because all reactives are flushed  and regenerated
    # with input changes but the renderUI has to make a round trip to the client
    # in get that input value.  Server to client to generate grades then back
    # again from client to server with user (or initial) grade selections
    if(input$att_level == 'hr')
      req(input$att_grades %in% ada_daily()$grade_level)

    # collect reactive values
    att_int <- att_date_range()
    daily <-  ada_daily()
    weekly <- ada_weekly()

    # filter by date
    daily2 <- daily %>% filter(date %within% att_int) %>% as.data.frame()
    weekly2 <- weekly %>% filter(date %within% att_int) %>% as.data.frame()

    enroll_attend_plot(daily2, weekly2)



    })


  students_filtered <- reactive({attend_student_ytd %>%
                         filter(grade_level == input$stu_grade)
                       })


  p_histo <- reactive(student_histogram(students_filtered() %>%
                        as.data.frame()))

  output$student_histogram <- renderPlot(p_histo())


  output$students <- DT::renderDataTable({
                        p_data <- p_histo()$data %>% as.data.frame()

                        x <- brushedPoints(p_data, input$plot_brush)

                        x <- x %>%
                          ungroup() %>%
                          mutate(
                            ada = ada/100,
                            grade_level = as.integer(grade_level)) %>%
                          select(
                              Student = lastfirst,
                              Grade = grade_level,
                               School = schoolabbreviation,
                               'Days Enrolled' = enrolled,
                               'Days Present' = present,
                               ADA = ada) %>%
                          arrange(desc(ADA))

                        DT::datatable(x,
                                      rownames = FALSE,
                                      style = "bootstrap",
                                      options =
                                        list(
                                          bPaginate = FALSE,
                                          dom = "t"
                                          )
                                      ) %>%
                          DT::formatPercentage("ADA", 1)
                        })

  output$att_goal_plot <- renderPlot({
    att_daily_goal_plot(
      attend_date_school,
      input$att_dates[[1]],
      input$att_dates[[2]],
      goals = goals,
      show_goals = input$show_ada)
    })


  output$leaders <- DT::renderDataTable({
                      attend_student_ytd %>% ungroup() %>%
                        filter(ada_rank <= .1,
                               schoolabbreviation == input$chronic_school) %>%
                        mutate(ada = ada/100,
                        grade_level = as.integer(grade_level)) %>%
                        select(Student = lastfirst,
                               Grade = grade_level,
                               School = schoolabbreviation,
                               'Days Enrolled' = enrolled,
                               'Days Present' = present,
                               ADA = ada) %>%
                        DT::datatable(
                          filter = "top",
                          rownames = FALSE,
                          options = list(
                            bPaginate = FALSE
                            )
                          ) %>%
                        DT::formatPercentage("ADA", 1)

                        })

```



```{r server_suspensions, include=FALSE}
todays_month<-factor(as.character(month(today(),
                                        label = TRUE)
                                  ),
                     levels= levels(susp_plot_data$Month),
                     ordered = TRUE)

susp_plot_data_2 <- susp_plot_data %>%
  filter(Type == "Suspension") %>%
  select(School, SY, Month, Type, N, Cum_N) %>%
  tidyr::complete(School, SY, Month, Type) %>%
  group_by(School, SY, Type) %>%
  mutate(Cum_N_2 = zoo::na.locf(Cum_N, na.rm=FALSE),
         Cum_N_4 = ifelse(is.na(Cum_N_2), 0, Cum_N_2)#,
         #Cum_N_4 = ifelse(SY=="SY17-18" &
         #                  as.integer(Month)>as.integer(todays_month),
         #                  NA,
         #                  Cum_N_3)
                          )

oss_combined <- susp_plot_data %>%
  select(school_name = School,
         SY,
         month = Month,
         N_susps = N,
         cum_susps = Cum_N,
         month_year = Month_Year) %>%
  mutate(school_name = as.character(school_name),
         month = forcats::fct_inorder(as.character(month))) %>%
  bind_rows(oss_rates %>%
              select(school_name,
                     month,
                     N_susps,
                     cum_susps, month_year, SY)
  ) %>%
  arrange(month_year) %>%
  mutate(month = factor(as.character(month),
                        levels = c("Aug", "Sep", "Oct", "Nov", "Dec", "Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul"),
                        ordered = T))




output$suspensions_viz <- renderPlot({
  p <- ggplot(susp_plot_data_2 %>%
                filter(SY %in% input$susp_years), aes(x=Month, y=Cum_N_4)) +
    geom_bar(aes(fill=SY),
             stat="identity", position="dodge") +
    facet_grid(.~School) +
    viridis::scale_fill_viridis(option = "D", discrete = TRUE, begin = 1, end = 0) +
    # scale_fill_manual("SY", values = c("#CFCCC1",
    #                               "#C49A6C",
    #                               "#17345B")) +
    #scale_alpha_discrete("SY", range = c(0.2, 1)) +
    theme_bw() +
    theme(axis.text.x = element_text(angle=45, hjust=1))

  p
  })
```



Daily Attendance and Enrollment {data-navmenu="Attendance"}
=======================================================================
Inputs {.sidebar}
-----------------------------------------------------------------------
```{r}
dateRangeInput(
              "att_dates",
              "Select dates:",
              start = last6weeks,
              end = thisweek,
              min = firstweek,
              max = thisweek,
              format = "mm-dd-yyyy"
             )

selectInput("att_level",
           "Select level:",
           choices = c("School" = "school",
                       "Grade" = "grade",
                       "Home Room" = "hr"),
           selected = "school")

conditionalPanel(condition = "input.att_level == 'grade' |
                  input.att_level == 'hr'",
                 selectInput("att_school",
                             "Select school:",
                             choices = schools,
                             selected = "KAMS")
                             )

conditionalPanel(condition = "input.att_level == 'hr'",
                 uiOutput("school_grades_ui")
                 )


```

Column
-----------------------------------------------------------------------

### Daily Attendance and Enrollment

```{r}
plotOutput("step_plot")
```

Daily ADA Goals {data-navmenu="Attendance"}
=======================================================================
Inputs {.sidebar}
-----------------------------------------------------------------------
```{r}
checkboxInput("show_ada",
              "Show ADA?",
              value = FALSE)       
```
This visual shows how far above or below KIPP Chicago's daily ADA each campus performed each school day over the last six weeks. Check the box above to add average daily attendance to each graphed pin.

Column
-----------------------------------------------------------------------

### Performance Above or Below Daily Attendance Goals
```{r}
plotOutput("att_goal_plot")
```


Student ADA {data-navmenu="Attendance"}
=======================================================================

Inputs {.sidebar}
-----------------------------------------------------------------------
You can click and drag in the histogram to the right in order to see
more detail on the students represented in each bar
```{r}
        selectInput("stu_grade",
         "Select Grade",
         choices  = 0:8,
         selected = 5)

```

Column
-----------------------------------------------------------------------

### Distribution of Individual Student Average Daily Attendance
```{r}
 plotOutput("student_histogram",
                      brush = brushOpts(
                                id = "plot_brush",
                                direction = "x")
                      )
```

### Detail table
```{r}
DT::dataTableOutput("students")
```



Most Chronic Absentees {data-navmenu="Attendance"}
=======================================================================
Inputs {.sidebar}
-----------------------------------------------------------------------
```{r}
selectInput("chronic_school",
         "Select School",
         choices = schools,
         selected = "KAMS",
         multiple = TRUE)
```
This table shows the worse attending 10% of students for each grade level at each school.

Column
-----------------------------------------------------------------------

### List of Chronically Absent Students
```{r}
DT::dataTableOutput("leaders")
```

Transfers
=======================================================================

Column {.tabset .tabset-fade}
-----------------------------------------------------------------------
### Cumulative Transfers to Date
```{r}
renderPlot({


    todays_month <- month(today(), label = TRUE, abbr = TRUE) %>%
      factor(levels = month_order, ordered = TRUE)

    transfer_scale <- c(scales::brewer_pal("qual", palette = 3)(nrow(transfer_reasons)-1),
                        "gray",
                        "lightgray"
                        )
    transfer_goals <- transfer_goals %>% ungroup() %>% mutate(sy = factor(sy, ordered = TRUE))

    ggplot(transfers_by_month_2 %>%
             dplyr::filter(!(month==todays_month &&
                      sy == "2017-2018")),
           aes(x = month, y=cum_transfers_2)) +
      geom_bar(aes(fill = reason, y=cum_transfers_2),
                position = "stack",
                stat = "identity") +
      geom_segment(data = transfer_goals,
                   aes(x=1, xend=12,
                       y=monthly_goal, yend= yearly_goal),
                   alpha = .6,
                   color = "purple") +
      facet_grid(sy ~ school_name) +
      scale_fill_manual(values = transfer_scale) +
      theme_light() +
      theme(axis.text.x = element_text(angle=45, hjust = 1)) +
      labs(fill = "Transfer\nReason",
           x = "Month",
           y = "Cumulative transfers\n(count)")
  })
```

### Historical Transfers by Month
```{r}
renderPlot({
  ggplot(transfers_by_month_2 %>%
         group_by(month, reason, school_name) %>%
         summarize(transfers = sum(transfers, na.rm = TRUE)),
       aes(x=month, y=reason)) +
  geom_tile(aes(fill = transfers),
            color = "black") +
  viridis::scale_fill_viridis(option = "D") +
  facet_grid(school_name ~ .) +
  theme_minimal() +
  labs(x = "", y="", fill = "Transfer Count")
})
```

Suspensions
=======================================================================
Inputs {.sidebar}
-----------------------------------------------------------------------
```{r}
selectInput("susp_years_2",
         "Select School Year",
         choices = unique(oss_combined$SY),
         selected = c("SY17-18"),
         multiple = TRUE)    
```

Column {.tabset}
-----------------------------------------------------------------------
```{r}
output$selected_oss <-  renderTable({

  #req(input$oss_click)
  validate(
    need(input$oss_click != "", "Click on a bar to see details for that school's and month's OSSs.")
  )

  school <- input$oss_click$panelvar1
  sy <- input$oss_click$panelvar2
  selected_month <- as.integer(round(input$oss_click$x))

  oss %>%
    mutate(int_month = as.integer(month),
           startdate = as.character(startdate)) %>%
   filter(school_name == school,
          as.integer(month) == selected_month,
          SY == sy) %>%  
    select('Last Name' = student_last,
           'First Name' = student_first,
           Grade = gradelevelshort,
           Date = startdate,
           School = school_name,
           Category = category,
           Summary = adminsummary
           ) %>%
    arrange(desc(Date))

})

```


### OSS Year to Date

```{r}

oss_select <- reactive({
  oss_combined %>%
         filter(SY %in% input$susp_years_2)
  })


output$oss <-renderPlot({
  ggplot(oss_select(),
         aes(x = month, y=cum_susps)) +
    geom_col() +
    geom_col(aes(y=N_susps), fill = "black", width=.5) +
    facet_grid(SY~school_name) +
    labs(x = "",
         y = "Suspensions (actual and cumulative)") +
    theme_linedraw()
  })

fillCol(flex = c(.5,.5),
  plotOutput("oss",
              click = "oss_click"),

  tableOutput("selected_oss")            
  )

```

### Current OSS Rates per 100 students

```{r}

oss_rates_cps <- tibble::tribble(
  ~school_name, ~period, ~susp_rate,
  "CPS", "Semester 1 (All)", 3.14,
  "CPS", "EOY (All)", 8.08,

  "CPS", "Semester 1 (AA)", 6.24,
  "CPS", "EOY (AA)", 16.09
)


renderPlot(
  oss_regional %>%
    filter(SY %in% input$susp_years_2) %>%
  ggplot(aes(x=school_name, y=cum_susp_rate)) +
    geom_col(aes(group = school_name, fill = regional), show.legend = FALSE) +
    geom_text(aes(label = round(cum_susp_rate,1)), vjust = 1, color = "white") +
    geom_hline(data=oss_rates_cps,
               aes(yintercept = susp_rate, color=period)) +
  scale_fill_manual(values = c("gray30", "gray10")) +              
  theme_linedraw()+
  labs(x = "",
       y = "Suspensions per 100 students")
       )
```

### ISS Year to Date
```{r}
output$selected_iss <-  renderTable({

  #req(input$oss_click)
  validate(
    need(input$iss_click != "", "Click on a bar to see details for that school's and month's OSSs.")
  )

  school_iss <- input$iss_click$panelvar1
  sy_iss <- input$iss_click$panelvar2
  selected_month_iss <- as.integer(round(input$iss_click$x))

  iss %>%
    mutate(int_month = as.integer(month),
          startdate = as.character(startdate)) %>%
   filter(school_name == school_iss,
          as.integer(month) == selected_month_iss) %>%  
    select('Last Name' = student_last,
           'First Name' = student_first,
           Grade = gradelevelshort,
           Date = startdate,
           School = school_name,
           Category = category,
           Summary = adminsummary
           ) %>%
    arrange(desc(Date))

})


output$iss <-renderPlot({
  ggplot(as.data.frame(iss_rates), aes(x = month, y=cum_susps)) +
  geom_col() +
  geom_col(aes(y=N_susps), fill = "black", width=.5) +
  #facet_grid(. ~ school_name) +
  facet_grid(. ~ school_name) +
  theme_linedraw() +
  labs(x = "",
       y = "In School Suspensions")
  })

fillCol(flex = c(.5,.5),
    plotOutput("iss",
               click = brushOpts(id = "iss_click")
               ),

    tableOutput("selected_iss")            
)

```


### Most Suspended Students
```{r}

output$days_hist <- renderPlot({
  oss %>%
    select(student_number,
           student_last,
           student_first,
           school_name,
           numdays) %>%
  ggplot(aes(x = as.integer(numdays))) +
  geom_histogram(bindwidth = 1) +
  scale_x_continuous(breaks = c(1:10)) +
  facet_grid(.~school_name) +
  theme_linedraw() +
  labs(title = 'Distribution of OSSs by days given',
       x = "# of days given",
       y = "# of students")
})


 output$sups_dt <-DT::renderDataTable({
    oss %>%
        group_by(student_number,
                 student_last,
                 student_first,
                 school_name,
                 gradelevelshort) %>%
        summarize(N = n(),
                  days_suspended  = sum(numdays)) %>%
         arrange(desc(days_suspended)) %>%
         ungroup() %>%
         dplyr::select(
           School = school_name,
           Grade = gradelevelshort,
           Last  = student_last,
           First  = student_first,
           Suspensions = N,
           'Days Suspended' = days_suspended
           )
 }, filter = 'top',
    extensions = c('Scroller', 'Buttons'),
    options = list(dom = 'Bfrtip',
                   scrollY = 400,
                   scroller = TRUE,
                   buttons = c('copy', 'csv', 'excel', 'pdf', 'print'))
    )

fillCol(flex = c(.25,.75),
        plotOutput("days_hist",
                   click = brushOpts(id = "iss_click")
                   ),

        DT::dataTableOutput("sups_dt")            
    )    
```


Teachers {data-navmenu="Observations"}
=======================================================================
Inputs {.sidebar}
-----------------------------------------------------------------------
```{r}
selectInput("obsv_schools",
         "Select School",
         choices = unique(teacher_last_two_weeks$school_abbrev),
         selected = "KAP",
         multiple = FALSE)    
```


Column {data-width=1000}
-----------------------------------------------------------------------

### YTD Observations of Teachers

```{r}
output$two_day_dt <- renderTable({
  validate(
    need(input$obsv_schools != "", "Select at least one school at left.")
  )

  teacher_last_two_weeks %>%
    filter(school_abbrev == input$obsv_schools) %>%
    mutate(two_weeks = n>=2) %>%
    select(Teacher = full_name,
           School = school_abbrev,
           "Obsv last two weeks" = n) #%>%


    #  DT::datatable()  %>%
    #  DT::formatStyle(1:3, valueColumns = 3,
    #    color = DT::styleInterval(
    #                         c(0,1),
    #                         c('red', '#8D8685', '#439539')
    #                     )
    #    )


    })



output$obs_teachers_plot <- renderPlot({

  max_month <- monthly_tb %>%
    ungroup() %>%
    summarize(max_month = max(as.integer(month)))

 obs_target <- max_month$max_month * 2


  teacher_ytd_tb %>%
    dplyr::filter(!is.na(topic_name)) %>%
    dplyr::filter(school_abbrev == input$obsv_schools) %>%
    dplyr::anti_join(admins %>% dplyr::select(user_id),
              by=c("topic_ids" = "user_id")) %>%
     dplyr::ungroup() %>%
     dplyr::arrange(n) %>%
     dplyr::mutate(name = forcats::fct_inorder(topic_name),
                   target_met = n >= obs_target) %>%
    ggplot(aes(x=name, y=n)) +
      geom_col(aes(fill = target_met),
               width = .75) +
      coord_flip() +
      viridis::scale_fill_viridis(discrete = TRUE,
                                  direction = -1) +
      theme_linedraw() +
      labs(y = "",
           x = "Observations to date",
           fill = sprintf("Target\n(%s obs = 2 x %s months)\nmet?",
                          obs_target,
                          max_month$max_month))
  })  

  plotOutput("obs_teachers_plot")

```

Column
-----------------------------------------------------------------------

### Observations over last two weeks (`r format(int_start(last_two_weeks), '%b %d')` to `r format(int_end(last_two_weeks), '%b %d')`)

```{r}
tableOutput("two_day_dt")
```

Admins {data-navmenu="Observations"}
=======================================================================
Inputs {.sidebar}
-----------------------------------------------------------------------
```{r}
selectInput("obsv_schools_admins",
         "Select School",
         choices = unique(teacher_last_two_weeks$school_abbrev),
         selected = "KAP",
         multiple = FALSE)    
```

Column
-----------------------------------------------------------------------

### Observations by admins by month

```{r}
output$obs_admins_plot <- renderPlot({
  monthly_tb %>%
  inner_join(admins %>% select(user_id), by=c("author_ids" = "user_id")) %>%
  filter(school_abbrev == input$obsv_schools_admins) %>%
  ggplot(aes(x=month, y=n)) +
    geom_col() +
    facet_wrap(~auth_name)  +
    theme_linedraw() +
    labs(x = "Month", y = "Observations conducted")
  })

plotOutput("obs_admins_plot")
```
