---
title: "School Culture"
output:
  flexdashboard::flex_dashboard:
    orientation: columns
    vertical_layout: fill
    theme: lumen
---

```{r setup, include=FALSE}
library(flexdashboard)
#library(Cairo)
library(lubridate)
library(ggplot2)
library(dplyr)
library(tidyr)

options(shiny.usecairo = TRUE)

load("/data/attendance.Rda")

load("/data/transfers.Rda")

schools <- unique(ada_weekly_school$schoolabbreviation)

grades <- list(KAP = 0:5,
               KAMS = 6:8,
               KCCP = 5:8,
               KBCP = 5:7)

month_order <- c("Oct", "Nov", "Dec", "Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sep")               

source('lib/attendance_helpers.R')

# ADA goals for each school
goals <- data_frame(schoolabbreviation =  c("KAP", "KAMS", "KCCP", "KBCP"),
                    goal = c(.955))

lastXweeks <- ymd(as.character(floor_date(today() - weeks(6), unit="week")+1))
last6weeks <- as.character(lastXweeks)
thisweek <- as.character(today())
firstweek <- as.character(floor_date(min(attend_date_school_grade$week_of_date)))


restart_time<-file.info('restart.txt')$mtime

update_time_stamp <- lubridate::stamp("Attendance data last updated on Tuesday, September 14, 2001 at 4:41 pm")(restart_time)

```

```{r server_attendance, include = FALSE}
att_date_range <- reactive(input$att_dates[[1]] %--% input$att_dates[[2]])



  # create dynamic panel
  output$school_grades_ui <- renderUI({

    school_grades <- ada_weekly_grade_hr %>%
      filter(schoolabbreviation == input$att_school) %>%
      select(grade_level)

    grades <- school_grades$grade_level %>% unique()

    selectInput("att_grades",
                 "Choose grade",
                 choices = as.character(grades),
                 selected = as.character(grades[1]),
                 multiple = FALSE)
    })

  # create daily data!
  ada_daily <- reactive({

    # Get data at school level
   if(input$att_level == "school") {


    daily_out <- attend_date_school %>%
      select(name = schoolabbreviation,
             date,
             week_in_sy,
             week_of_date,
             week_of_date_short_label,
             enrolled,
             present) %>%
      mutate(goal = round(.97*enrolled),
             day = wday(date)) %>%
      gather(key = "variable",
             value = "value", enrolled:goal)

   }

    # Munge data for Grade level
    # To do: create menu item for school
    if(input$att_level == "grade") {

      daily_out <- attend_date_grade %>%
        filter(schoolabbreviation == input$att_school) %>%
        select(name = grade_level,
               schoolabbreviation,
               date,
               week_in_sy,
               week_of_date,
               week_of_date_short_label,
               enrolled,
               present) %>%
        mutate(goal = round(.97*enrolled),
               day = wday(date)) %>%
        gather(key = "variable",
               value = "value", enrolled:goal)

    }

    if(input$att_level == "hr") {

      daily_out <- attend_date_grade_hr %>%
        filter(grade_level == as.numeric(input$att_grades),
               schoolabbreviation == input$att_school
               ) %>%
        select(name = home_room,
               grade_level,
               schoolabbreviation,
               date,
               week_in_sy,
               week_of_date,
               week_of_date_short_label,
               enrolled,
               present) %>%
        mutate(goal = round(.97*enrolled),
               day = wday(date)) %>%
        gather(key = "variable",
               value = "value", enrolled:goal)

    }

  # return
  daily_out

  })







  # ada_weekly reactive
  ada_weekly <- reactive({

    # if school
    if(input$att_level == "school") {
      weekly_out <- ada_weekly_school %>%
        mutate(name = schoolabbreviation,
               threshold = ifelse(weekly_ada>=.97*100,
                                  sprintf("Weekly ≥ %s%%", round(.97*100)),
                                  sprintf("Weekly < %s%%",  round(.97*100))
                                  )
              )
     }



    # if grade
    if(input$att_level == "grade") {

      weekly_out <- ada_weekly_school_grade %>%
      filter(schoolabbreviation == input$att_school) %>%
        mutate(name = grade_level,
               threshold = ifelse(weekly_ada>=.97*100,
                                  sprintf("Weekly ≥ %s%%", round(.97*100)),
                                  sprintf("Weekly < %s%%",  round(.97*100))
                                  )
              )
     }

    # if home room
    if(input$att_level == "hr") {


      weekly_out <- ada_weekly_grade_hr %>%
        filter(grade_level == as.numeric(input$att_grades),
               schoolabbreviation == input$att_school
               ) %>%
        mutate(name = home_room,
               threshold = ifelse(weekly_ada>=.97*100,
                                  sprintf("Weekly ≥ %s%%", round(.97*100)),
                                  sprintf("Weekly < %s%%",  round(.97*100))
                                  )
              )
   }

  #return
  weekly_out

  })






  # Step plot of enrollment and attendance
  output$step_plot <- renderPlot({

    # verify that initial grade level exists for school if att_level == 'hr'
    # you need to do this because all reactives are flushed  and regenerated
    # with input changes but the renderUI has to make a round trip to the client
    # in get that input value.  Server to client to generate grades then back
    # again from client to server with user (or initial) grade selections
    if(input$att_level == 'hr')
      req(input$att_grades %in% ada_daily()$grade_level)

    # collect reactive values
    att_int <- att_date_range()
    daily <-  ada_daily()
    weekly <- ada_weekly()

    # filter by date
    daily2 <- daily %>% filter(date %within% att_int)
    weekly2 <- weekly %>% filter(date %within% att_int)

    enroll_attend_plot(daily2, weekly2)



    })


  students_filtered <- reactive({attend_student_ytd %>%
                         filter(grade_level == input$stu_grade)
                       })


  p_histo <- reactive(student_histogram(students_filtered()))

  output$student_histogram <- renderPlot(p_histo())


  output$students <- DT::renderDataTable({
                        p_data <- p_histo()$data

                        x <- brushedPoints(p_data, input$plot_brush)

                        x <- x %>%
                          ungroup() %>%
                          mutate(
                            ada = ada/100,
                            grade_level = as.integer(grade_level)) %>%
                          select(
                              Student = lastfirst,
                              Grade = grade_level,
                               School = schoolabbreviation,
                               'Days Enrolled' = enrolled,
                               'Days Present' = present,
                               ADA = ada) %>%
                          arrange(desc(ADA))

                        DT::datatable(x,
                                      rownames = FALSE,
                                      style = "bootstrap",
                                      options =
                                        list(
                                          bPaginate = FALSE,
                                          dom = "t"
                                          )
                                      ) %>%
                          DT::formatPercentage("ADA", 1)
                        })

  output$att_goal_plot <- renderPlot({
    att_daily_goal_plot(
      attend_date_school,
      input$att_dates[[1]],
      input$att_dates[[2]],
      goals = goals,
      show_goals = input$show_ada)
    })


  output$leaders <- DT::renderDataTable({
                      attend_student_ytd %>% ungroup() %>%
                        filter(ada_rank <= .1,
                               schoolabbreviation == input$chronic_school) %>%
                        mutate(ada = ada/100,
                        grade_level = as.integer(grade_level)) %>%
                        select(Student = lastfirst,
                               Grade = grade_level,
                               School = schoolabbreviation,
                               'Days Enrolled' = enrolled,
                               'Days Present' = present,
                               ADA = ada) %>%
                        DT::datatable(
                          filter = "top",
                          rownames = FALSE,
                          options = list(
                            bPaginate = FALSE
                            )
                          ) %>%
                        DT::formatPercentage("ADA", 1)

                        })

```


```{r server_transfers, include = FALSE}
output$transfer_plot <- renderPlot({


    todays_month <- month(today(), label = TRUE, abbr = TRUE) %>%
      factor(levels = month_order, ordered = TRUE)

    transfer_scale <- c(scales::brewer_pal("qual", palette = 3)(nrow(transfer_reasons)-3),
                        "gray",
                        "lightgray"
                        )


    ggplot(transfers_by_month_2 %>%
             filter(!(month > todays_month &
                      sy == "2015-2016")
                    ),
           aes(x = month, y=cum_transfers_2)) +
      geom_bar(aes(fill = factor(reason), y=cum_transfers_2),
                position = "stack",
                stat = "identity") +
      geom_segment(data = transfer_goals,
                   aes(x=1, xend=12,
                       y=monthly_goal, yend= yearly_goal),
                   alpha = .6,
                   color = "purple") +
      facet_grid(sy ~ school_name) +
      scale_fill_manual(values = transfer_scale) +
      theme_light() +
      theme(axis.text.x = element_text(angle=45, hjust = 1)) +
      labs(fill = "Transfer\nReason",
           x = "Month",
           y = "Cumulative transfers\n(count)")

    })


```


Daily Attendance and Enrollment {data-navmenu="Attendance"}
=======================================================================
Inputs {.sidebar}
-----------------------------------------------------------------------
```{r}
dateRangeInput(
              "att_dates",
              "Select dates:",
              start = last6weeks,
              end = thisweek,
              min = firstweek,
              max = thisweek,
              format = "mm-dd-yyyy"
             )

selectInput("att_level",
           "Select level:",
           choices = c("School" = "school",
                       "Grade" = "grade",
                       "Home Room" = "hr"),
           selected = "School")

conditionalPanel(condition = "input.att_level == 'grade' |
                  input.att_level == 'hr'",
                 selectInput("att_school",
                             "Select school:",
                             choices = schools,
                             selected = "KAMS")
                             )

conditionalPanel(condition = "input.att_level == 'hr'",
                 uiOutput("school_grades_ui")
                 )


```

Column
-----------------------------------------------------------------------

### Daily Attendance and Enrollment

```{r}
plotOutput("step_plot")
```

Daily ADA Goals {data-navmenu="Attendance"}
=======================================================================
Inputs {.sidebar}
-----------------------------------------------------------------------
```{r}
checkboxInput("show_ada",
              "Show ADA?",
              value = FALSE)       
```
This visual shows how far above or below KIPP Chicago's daily ADA each campus performed each school day over the last six weeks. Check the box above to add average daily attendance to each graphed pin.

Column
-----------------------------------------------------------------------

### Performance Above or Below Daily Attendance Goals
```{r}
plotOutput("att_goal_plot")
```


Student ADA {data-navmenu="Attendance"}
=======================================================================

Inputs {.sidebar}
-----------------------------------------------------------------------
You can click and drag in the histogram to the right in order to see
more detail on the students represented in each bar
```{r}
        selectInput("stu_grade",
         "Select Grade",
         choices  = 0:8,
         selected = 5)

```

Column
-----------------------------------------------------------------------

### Distribution of Inidividual Student Average Daily Attendance
```{r}
 plotOutput("student_histogram",
                      brush = brushOpts(
                                id = "plot_brush",
                                direction = "x")
                      )
```

### Detail table
```{r}
DT::dataTableOutput("students")
```



Most Chronic Absentees {data-navmenu="Attendance"}
=======================================================================
Inputs {.sidebar}
-----------------------------------------------------------------------
```{r}
selectInput("chronic_school",
         "Select School",
         choices = schools,
         selected = "KAMS",
         multiple = TRUE)
```
This table shows the worse attending 10% of students for each grade level at each school.

Column
-----------------------------------------------------------------------

### List of Chronically Absent Students
```{r}
DT::dataTableOutput("leaders")
```

Transfers
=======================================================================

Column {.tabset}
-----------------------------------------------------------------------
### Cumulative Transfers to Date
```{r}
plotOutput("transfer_plot")
```

### Historical Transfers by Month
```{r}
renderPlot({
  ggplot(transfers_by_month_2 %>%
         group_by(month, reason, school_name) %>%
         summarize(transfers = sum(transfers, na.rm = TRUE)),
       aes(x=month, y=reason)) +
  geom_tile(aes(fill = transfers),
            color = "black") +
  viridis::scale_fill_viridis(option = "D") +
  facet_grid(school_name ~ .) +
  theme_minimal() +
  labs(x = "", y="", fill = "Transfer Count")
})
```

Suspensions
=======================================================================
### Coming Soon!
